Curve counting in genus one: elliptic singularities {\it \&} relative geometry

### 2019-06-28 EllipticSingularitiesSplitting.tex
spell check

\begin{proof}
We prove the results via the forgetful morphism
$$
\nu: \VZ_{1,\alpha}(\mathbb P^N|H,d)\to \VZ_{1,n}(\mathbb P^N,d),
$$
which remembers only the stabilization of the collapsed map. The map is certainly centrally aligned, and we have already argued above that it satisfies the factorisation property for $\mathbb P^N$ after composition. The moduli space $\VZ_{1,\alpha}(\mathbb P^N|H,d)$ is proper. Indeed, the verification of the valuative criterion follows exactly as in Vakil's Lemma~\cite[Lemma~5.9]{Vre}. A direct proof that the factorisation property is a closed condition may be found in~\cite[Theorem~4.3]{RSPW}.

@@

Assume now that the elliptic curve maps into higher level. Note that the relative logarithmic tangent bundle of the expansion, over the base $\mathbb P^N$ is a trivial line bundle of rank $1$, as the fibres are toric. Consider an $S$-family of maps. After replacing the source $C$ by a destabilisation, we have maps
\bcd
C \ar[r] & \overline{C}_F \ar[r] \ar[d] & \overline{C}_B \ar[d] \\
& \mathbb P^N[s] \ar[r] & \mathbb P^N
\ecd
factorizing completely. Let $f_B$ and $f$ be the maps from $C$ to the base and total space respectively. Examining the morphism $\nu$, we see that there is a map
\[
\mathrm{Def}(C,f_B)\to \mathrm{Obs}(f/(C,f_B)) = H^1(\overline{C}_F,\mathcal O_{\overline{C}_F}),
\]
where the former is the space of deformations of the curve and map to the base, as a factorised centrally aligned map. The latter is the space of obstructions to lifting a map to the base into the total space. The cokernel of this map are the absolute obstructions, which we will show vanishes, proving logarithmic smoothness.

@@

\noindent The logarithmic smoothness established in Theorem~\ref{thm: log-smoothness} implies that $\VZ_{1,\alpha}(\mathbb P^N|H,d)$ is an orbifold toroidal embeddeding. Consequently, the irreducible components of its boundary and their intersections form a stratification of the space. This stratification will be important in the sequel.

@@

\begin{definition}
Let $\plC$ be a tropical curve of genus one and let $\plC_0$ be its minimal subcurve of genus one. A tropical map $F: \plC\to \mathbb R_{\geq 0}$ is said to be \textbf{well-spaced} if one of the following two conditions are satisfied:
\begin{enumerate}
    \item no open neighborhood of $\plC_0$ is contracted to a point in $\mathbb R_{>0}$, or
    \item if an open neighborhood of $\plC_0$ is contracted and $t_1,\ldots,t_k$ are the flags whose vertex is mapped to $F(\plC_0)$, but along which $F$ has nonzero slope; then, the minimum of the distances from $\plC_0$ to $t_i$ occurs for at least two indices $i$.
\end{enumerate}
\end{definition}

@@

\begin{proof}
The required result is essentially contained in~\cite[Section 4]{RSPW2}, so we explain how to deduce the requisite result from the one in loc. cit. We may focus on a single component of the expansion $\mathbb P^N[s]$ that contains the image of a contracted genus one subcurve, as this is the only relevant case. We may thus replace the target with the projective bundle $\mathbb P(\mathcal O(1)\oplus \mathcal O)$ over $\mathbb P^{N-1}$ equipped with the divisorial logarithmic structure from the $0$ and $\infty$ sections. Let $p$ be the point to which the genus one subcurve is contracted. Passing to an open neighborhood of $p$, the map to the bundle is given by a rational function $f$ on an open curve $C^\circ$, determined by the the subgraph formed by $\plC_0$ and the flags $t_i$ described in the definition above. To describe the tropical map to $\mathbb R_{>0}$, we observe that $\plC_0$ is contracted to a fixed point $q\in\mathbb R_{>0}$. The flags at $t_i$ correspond to nodes or markings of $C$. The pole orders of $f$ at these distinguished points determine the slopes of the tropical map. We are now exactly in the situation considered in~\cite[Second Paragraph of \S~4.6]{RSPW2}, and Lemma~4.6.1 of loc. cit. guarantees the well-spacedness as required.
\end{proof}

@@

\noindent
\textbf{Step 3.} Given a tropical map $F: \plC\to \mathbb R_{\geq 0}$ parameterized by $\Sigma^{\mathrm{Kim}}$, there is a largest radius $\delta$ (possibly equal to $0$) such that every vertex strictly contained in the circle of radius $\delta$ around the core has degree-marking $d=0$. Let $
\Sigma^{\mathrm{cen}}$ be the subdivision obtained by requiring that the vertices contained within the circle of radius $\delta$ around the core of $\plC$ are totally ordered. This involves introducing cones along which certain $\Z$-linear combinations of edge lengths are identified.\\

@@

\begin{proof}
The construction above has been given to mimic the construction of the space $
\VZ_{1,\alpha}(\mathbb P^N|H,d)$. Specifically, the fact that the cone complex $\Sigma^{\mathrm{cen}}$ is the cone complex attached to the logarithmic stack $\widetilde{\VZ}_{1,\alpha}(\mathbb P^N|H,d)$ of centrally aligned maps to expansions, follows immediately from its description as a subcategory of the fibreed category (over logarithmic schemes) of $\mathcal M^{\mathrm{log}}_{1,\alpha}(\mathbb P^N|H,d)$. To complete the result, we note that $
\VZ_{1,\alpha}(\mathbb P^N|H,d)$ has a strict map to $
\VZ_{1,\alpha}(\mathbb P^N|H,d)$, so its cone complex is a subcomplex of $\Sigma^{\mathrm{cen}}$. The fact that it must be contained in the subcomplex of well-spaced curves is immediate from Proposition~\ref{prop: well-spaced}.
\end{proof}

\subsection{Indexing the strata} \label{subsection indexing strata} The dimension-$k$ cones in $\Sigma_{1,\alpha}(\PP^N|H,d)$ can be enumerated as follows. First, the cones in $\Sigma=\Sigma^{\Kim}$ are indexed by \textbf{combinatorial types} of tropical maps to $\RR_{\geq 0}$. Here a combinatorial type encodes all of the data of a tropical map, except for the edge lengths. To be more precise, a combinatorial type $\Delta$ consists of:
\begin{enumerate}
\item a finite graph $\sqC$ and topological (not metric) subdivision $\widetilde\RR_{\geq 0}$ of $\RR_{\geq 0}$;
\item genus, degree and marking assignments on the vertices;
\item the data of which vertex of $\widetilde\RR_{\geq 0}$ each vertex of $\sqC$ is mapped to;
\item integral slopes along the edges (both finite and infinite).
\end{enumerate}
The corresponding cone $\sigma \in \Sigma^{\operatorname{Kim}}$ is then given by the resulting moduli space of tropical maps, given by choices of edge lengths on source and target which produce a continuous tropical map. Given $\sigma\in \Sigma^{\operatorname{Kim}}$ we then produce all the cones in $\Sigma_{1,\alpha}(\PP^N|H,d)$ mapping to $\sigma$ by performing Steps 3--4 outlined above. This amounts to taking a particular subdivision of $\sigma$. Note that in this process, new cones are created which map into larger-dimensional cones of $\Sigma^{\operatorname{Kim}}$. The process for enumerating the codimension $k$ strata of $\VZ_{1,\alpha}(\PP^N|H,d)$ is therefore:
\begin{enumerate}
\item fix a combinatorial type $\Delta$;
\item perform the subdvidision of the resulting cone $\sigma \in \Sigma^{\operatorname{Kim}}$;
\item identify the dimension $k$ cones of that subdivision.
\end{enumerate}
In \S \ref{section reduced splitting} we will perform this analysis in the case $k=1$.

@@

\begin{proof} The locus where the logarithmic structure is trivial coincides precisely with the  locus where the source curve is smooth, and consequently not mapped inside $H$. Since $\VZ_{1,\alpha}(\PP^N|H,d)$ is toroidal this locus is open and dense. By definition $\Dcal_{1,\alpha,k}(\PP^N|H,d)$ is contained in the  complement of this locus and can be expressed as a union of strata of positive codimension. Since $\Dcal_{1,\alpha,k}(\PP^N|H,d)$ has codmension $1$, it must therefore be equal to a union of logarithmic divisors.\end{proof}

In \S \ref{subsection indexing strata} we discussed a procedure for enumerating the logarithmic strata of $\VZ_{1,\alpha}(\PP^N|H,d)$, using the cones of the tropicalization $\Sigma_{1,\alpha}(\PP^N|H,d)$. Every divisorial logarithmic stratum is obtained as follows:
\begin{enumerate}
\item choose a combinatorial type $\Delta$ of a tropical map;
\item subdivide the corresponding tropical moduli space $\sigma$ to align the type;
\item choose a ray in this subdivision.
\end{enumerate}
This process contains some redundancies: upon choosing a ray in the tropical moduli space, some of the edge lengths or vertex positions may get set to $0$. This induces a generization of the intial combinatorial type $\Delta$, given by contracting the corresponding edges of the dual graph and moving the corresponding vertices from $\RR_{>0}$ to $0$. When we speak of the combinatorial type of a stratum in $\VZ_{1,\alpha}(\PP^N|H,d)$, we will always mean this generisation. This is independent of the choice of $\Delta$, and in fact we can and will always choose $\Delta$ to coincide with the generization.

@@

The map $\Dcal\to\Ecal$ is an isomorphism away from the exceptional centers, and the latter is contained in the locus where the core is contracted. Given an element of $\Dcal$ we can split it along the nodes $q_1,\ldots,q_r$. It is then clear that $\sqC_1$ is aligned. We claim that $\sqC$ satisfies the factorisation property if and only if $\sqC_1$ does. This immediately implies the lemma.

@@

We summarize that the contribution of each term $\lambda_\Dcal [\Dcal]$ is given by integrating over the appropriate fibre product and then multiplying the result by $\lambda_\Dcal \cdot \deg\rho = \prod_{i=1}^r m_i$.

@@

For each vertex of the target there must always be at least one vertex in the preimage which lies on the circle. This is implied by the fact that the cone $\tau$ corresponding to $\Dcal$ is a ray. If the vertex on the circle is strictly semistable -- so that in particular the collapsed map $f_B$ is constant along the corresponing component -- it can be seen that the factorisation condition is unaffected by $\Gm$-scaling the coordinate corresponding to that vertex in the moduli space of attaching data, so the map $\Dcal\to\Ecal$ has generic fibre of positive dimension and the contribution vanishes.

@@

Finally, in case the maximal contact order is $m=1$, the image tangent vectors must be linearly dependent, in which case we can deduce as above that factorisation will be satisfied indepedently of the choice of alignment.
\end{proof}
For the remainder of this section we use the following notation for the combinatorial type $\Delta$; we let $\sqC_0$ denotes the vertex of dual graph corresponding to the core, $\sqC_1,\ldots,\sqC_r$ the other stable vertices (genus zero and lying on the circle, by the previous lemma) and $q_1,\ldots,q_r$ the corresponding splitting nodes.

@@

The combinatorics of the flattening can be described explicitly. Consider the map on tropicalizations
\[
\widetilde{\Dcal}^{\operatorname{trop}} \to \Ecal^{\operatorname{trop}}.
\]
The tropical curves on the left inherit a central alignment from the moduli space $\widetilde{\VZ}_{1,\alpha}(\PP^N|H,d)$, which orders a certain subset of the edge lengths. The polyhedral criterion for equidimensionality is that every cone of the source maps surjectively onto a cone of the target. There is a universal logarithmic modification $\widetilde\Ecal \to \Ecal$ such that $\widetilde\Dcal \to \widetilde\Ecal$ is flat, constructed by taking the minimal subdivison of $\Ecal^{\operatorname{trop}}$ such that cones map surjectively to cones.

@@

\begin{lemma}
 For every $\Lambda_B\subseteq \{\sqC_1,\ldots,\sqC_r\}$ and $\Lambda_F=\{\sqC_1,\ldots,\sqC_r\}\setminus \Lambda_B$, $s$ vanishes along the loci where the factors corresponding to $\Lambda_B$ in $\widetilde \Ecal$ degenerate so that $f_i\colon C_i\to\PP^N$ is constant in a neighbourhood of $q_i$, and the fibre coordinates $(\lambda_j)_{j\in\Lambda_F}$ along $p\colon\mathcal P\to\widetilde \Ecal$ are set to be $0$.
\end{lemma}
Call these loci $\Xi_\Lambda$ for $\Lambda=(\Lambda_B,\Lambda_F)$ (compare with \cite[\S 3.2]{VZ} for an analogous phenomenon in the non-relative case); clearly, they may be in excess with respect to the expected codimension of $V(s)$, that is $N-1$. These loci are logarithmic substrata; therefore, there exists a logarithmic blow-up that principalizes them. We claim which $\widetilde\Dcal$ is (possibly a refinement of) such a blow-up. 
\begin{lemma}
 The pullback of $\Xi_\Lambda$ is principal in $\widetilde\Dcal$.
\end{lemma}
\begin{proof}We analyse the ideal of $\Xi_\Lambda$: it is generated by
 \begin{enumerate}
  \item equations on $\widetilde\Ecal$ cutting the locus where $C_i$ degenerates so that $f_i$ is constant on a neighbourhood of $q_i$ for $i\in \Lambda_B$: these correspond to the tropical functions $\delta_i-\lambda(q_i)$ where $\lambda$ is the distance function to the core;
  \item equations on the fibre of $p$ for the linear subspace corresponding to $i\in \Lambda_F$: tropically, these correspond to the $\ell_i$ discussed in Lemma \ref{lem:generic_proj_bundle}.
 \end{enumerate}
By construction, on $\widetilde \Dcal$ it is always possible to tell which of these tropical lengths is the shortest (the circle we have drawn on $\Xi_\Lambda$ is degenerate -- it passes through no vertex of positive horizontal degree -- therefore we need to decide for a strictly larger circle in order to lift to $\widetilde D$), which corresponds to principalising the ideal (see Figure \ref{fig:principalisation}).
\end{proof}

@@

%Draw ldots
\draw [color=blue] (4,0.5) node[right]{$\ldots$};
\end{tikzpicture}
\end{center}
\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
with all other marked points contained on $\sqC_1$. This is isomorphic to $\VZ_{1,\alpha}(X|Y,\beta)$ and when we restrict $\tilde\gamma$ to this locus we obtain precisely the class $\gamma$ which we started with. Thus the left-hand side of \eqref{step 1 recursion} may be written as
\begin{equation*} (\alpha_1-1) I + \fgt_y^\st \left( (\alpha_1-1)\psi_1 + \ev_1^\st Y\right)  \tilde\gamma \cap [\VZ_{1,(\alpha-e_1)\cup(1)}(X|Y,\beta)]\end{equation*}
where $I$ is the invariant we are trying to compute. The second term is a forgetful relative invariant with the same degree and number of true markings (since $y$ is fictitious), and strictly smaller true tangency; hence it is recursively known. We now examine the right-hand side of \eqref{step 1 recursion}. Recall that all of the genus zero data has already been computed, so we only need to focus on the genus one pieces. First consider the type I loci. The genus one piece has strictly smaller degree (and hence is recursively known) exept in the following situation (with some stable distribution of the remaining markings):
\begin{center}
\begin{tikzpicture}[scale=1]
\draw (0,0) to (2,0);
\draw [fill=white] (0,0) circle[radius=3pt];
\draw (0,-0.1) node[below]{\small$d$};
\draw (0,0) node[above]{\small$\sqC_1$};
\draw [fill=black] (2,0) circle[radius=3pt];
\draw (2,-0.1) node[below]{\small$0$};
\draw (2,0) node[above]{\small$\sqC_0$};
\draw [->] (2,0) -- (3,0);
\draw (2.9,0) node[right]{$x_1$};

@@

We will make use of a recursion formula for rubber spaces analogous to the recursion formula for relative spaces used in Steps 1--3. Following \cite{EKatz}, there is a line bundle $L_y^{\not\in \operatorname{bot}}$ on $\VZ^{\leftrightarrow}_{1,\alpha\cup (0)}(\mathbb P|Y_0+Y_\infty,\beta)$, together with a section $s_y^{\not\in\operatorname{bot}}$ whose vanishing locus consists of the locus $\Dcal(y)$ of rubber maps where $y$ is not mapped into the bottom level of the expanded target. As in \S \ref{Section Gathmann line bundle}, we can give a logarithmic interpretation of this: it corresponds to the piecewise-linear function on the tropical moduli space which associates, to every rubber tropical map, the distance between $\varphi(\sqC_y)$ and the leftmost vertex of the tropical target, where $\sqC_y$ is the vertex of the source curve containing the flag corresponding to $y$. Using this tropical description, we may easily calculate the vanshing orders of $s_y^{\not\in\operatorname{bot}}$ along the various components of $\Dcal(y)$, and show that $\cchern_1(L_y^{\not\in\operatorname{bot}}) = \Psi_0 - \ev_y^\st Y$ (see also \cite{EKatzLB}, where similar results in the non-reduced setting are obtained, using different methods). From this, we obtain a rubber recursion formula

>>>>

\begin{proof}
We prove the results via the forgetful morphism
$$
\nu: \VZ_{1,\alpha}(\mathbb P^N|H,d)\to \VZ_{1,n}(\mathbb P^N,d),
$$
which remembers only the stabilisation of the collapsed map. The map is certainly centrally aligned, and we have already argued above that it satisfies the factorisation property for $\mathbb P^N$ after composition. The moduli space $\VZ_{1,\alpha}(\mathbb P^N|H,d)$ is proper. Indeed, the verification of the valuative criterion follows exactly as in Vakil's Lemma~\cite[Lemma~5.9]{Vre}. A direct proof that the factorisation property is a closed condition may be found in~\cite[Theorem~4.3]{RSPW}.

@@

Assume now that the elliptic curve maps into higher level. Note that the relative logarithmic tangent bundle of the expansion, over the base $\mathbb P^N$ is a trivial line bundle of rank $1$, as the fibres are toric. Consider an $S$-family of maps. After replacing the source $C$ by a destabilisation, we have maps
\bcd
C \ar[r] & \overline{C}_F \ar[r] \ar[d] & \overline{C}_B \ar[d] \\
& \mathbb P^N[s] \ar[r] & \mathbb P^N
\ecd
factorising completely. Let $f_B$ and $f$ be the maps from $C$ to the base and total space respectively. Examining the morphism $\nu$, we see that there is a map
\[
\mathrm{Def}(C,f_B)\to \mathrm{Obs}(f/(C,f_B)) = H^1(\overline{C}_F,\mathcal O_{\overline{C}_F}),
\]
where the former is the space of deformations of the curve and map to the base, as a factorised centrally aligned map. The latter is the space of obstructions to lifting a map to the base into the total space. The cokernel of this map are the absolute obstructions, which we will show vanishes, proving logarithmic smoothness.

@@

\noindent The logarithmic smoothness established in Theorem~\ref{thm: log-smoothness} implies that $\VZ_{1,\alpha}(\mathbb P^N|H,d)$ is an orbifold toroidal embedding. Consequently, the irreducible components of its boundary and their intersections form a stratification of the space. This stratification will be important in the sequel.

@@

\begin{definition}
Let $\plC$ be a tropical curve of genus one and let $\plC_0$ be its minimal subcurve of genus one. A tropical map $F: \plC\to \mathbb R_{\geq 0}$ is said to be \textbf{well-spaced} if one of the following two conditions are satisfied:
\begin{enumerate}
    \item no open neighbourhood of $\plC_0$ is contracted to a point in $\mathbb R_{>0}$, or
    \item if an open neighbourhood of $\plC_0$ is contracted and $t_1,\ldots,t_k$ are the flags whose vertex is mapped to $F(\plC_0)$, but along which $F$ has nonzero slope; then, the minimum of the distances from $\plC_0$ to $t_i$ occurs for at least two indices $i$.
\end{enumerate}
\end{definition}

@@

\begin{proof}
The required result is essentially contained in~\cite[Section 4]{RSPW2}, so we explain how to deduce the requisite result from the one in loc. cit. We may focus on a single component of the expansion $\mathbb P^N[s]$ that contains the image of a contracted genus one subcurve, as this is the only relevant case. We may thus replace the target with the projective bundle $\mathbb P(\mathcal O(1)\oplus \mathcal O)$ over $\mathbb P^{N-1}$ equipped with the divisorial logarithmic structure from the $0$ and $\infty$ sections. Let $p$ be the point to which the genus one subcurve is contracted. Passing to an open neighbourhood of $p$, the map to the bundle is given by a rational function $f$ on an open curve $C^\circ$, determined by the the subgraph formed by $\plC_0$ and the flags $t_i$ described in the definition above. To describe the tropical map to $\mathbb R_{>0}$, we observe that $\plC_0$ is contracted to a fixed point $q\in\mathbb R_{>0}$. The flags at $t_i$ correspond to nodes or markings of $C$. The pole orders of $f$ at these distinguished points determine the slopes of the tropical map. We are now exactly in the situation considered in~\cite[Second Paragraph of \S~4.6]{RSPW2}, and Lemma~4.6.1 of loc. cit. guarantees the well-spacedness as required.
\end{proof}

@@

\noindent
\textbf{Step 3.} Given a tropical map $F: \plC\to \mathbb R_{\geq 0}$ parameterised by $\Sigma^{\mathrm{Kim}}$, there is a largest radius $\delta$ (possibly equal to $0$) such that every vertex strictly contained in the circle of radius $\delta$ around the core has degree-marking $d=0$. Let $
\Sigma^{\mathrm{cen}}$ be the subdivision obtained by requiring that the vertices contained within the circle of radius $\delta$ around the core of $\plC$ are totally ordered. This involves introducing cones along which certain $\Z$-linear combinations of edge lengths are identified.\\

@@

\begin{proof}
The construction above has been given to mimic the construction of the space $
\VZ_{1,\alpha}(\mathbb P^N|H,d)$. Specifically, the fact that the cone complex $\Sigma^{\mathrm{cen}}$ is the cone complex attached to the logarithmic stack $\widetilde{\VZ}_{1,\alpha}(\mathbb P^N|H,d)$ of centrally aligned maps to expansions, follows immediately from its description as a subcategory of the fibred category (over logarithmic schemes) of $\mathcal M^{\mathrm{log}}_{1,\alpha}(\mathbb P^N|H,d)$. To complete the result, we note that $
\VZ_{1,\alpha}(\mathbb P^N|H,d)$ has a strict map to $
\VZ_{1,\alpha}(\mathbb P^N|H,d)$, so its cone complex is a subcomplex of $\Sigma^{\mathrm{cen}}$. The fact that it must be contained in the subcomplex of well-spaced curves is immediate from Proposition~\ref{prop: well-spaced}.
\end{proof}

\subsection{Indexing the strata} \label{subsection indexing strata} The dimension-$k$ cones in $\Sigma_{1,\alpha}(\PP^N|H,d)$ can be enumerated as follows. First, the cones in $\Sigma=\Sigma^{\Kim}$ are indexed by \textbf{combinatorial types} of tropical maps to $\RR_{\geq 0}$. Here a combinatorial type encodes all of the data of a tropical map, except for the edge lengths. To be more precise, a combinatorial type $\Delta$ consists of:
\begin{enumerate}
\item a finite graph $\sqC$ and topological (not metric) subdivision $\widetilde\RR_{\geq 0}$ of $\RR_{\geq 0}$;
\item genus, degree and marking assignments on the vertices;
\item the data of which vertex of $\widetilde\RR_{\geq 0}$ each vertex of $\sqC$ is mapped to;
\item integral slopes along the edges (both finite and infinite).
\end{enumerate}
The corresponding cone $\sigma \in \Sigma^{\operatorname{Kim}}$ is then given by the resulting moduli space of tropical maps, given by choices of edge lengths on source and target which produce a continuous tropical map. Given $\sigma\in \Sigma^{\operatorname{Kim}}$ we then produce all the cones in $\Sigma_{1,\alpha}(\PP^N|H,d)$ mapping to $\sigma$ by performing Steps 3--4 outlined above. This amounts to taking a particular subdivision of $\sigma$. Note that in this process, new cones are created which map into larger-dimensional cones of $\Sigma^{\operatorname{Kim}}$. The process for enumerating the codimension $k$ strata of $\VZ_{1,\alpha}(\PP^N|H,d)$ is therefore:
\begin{enumerate}
\item fix a combinatorial type $\Delta$;
\item perform the subdivision of the resulting cone $\sigma \in \Sigma^{\operatorname{Kim}}$;
\item identify the dimension $k$ cones of that subdivision.
\end{enumerate}
In \S \ref{section reduced splitting} we will perform this analysis in the case $k=1$.

@@

\begin{proof} The locus where the logarithmic structure is trivial coincides precisely with the  locus where the source curve is smooth, and consequently not mapped inside $H$. Since $\VZ_{1,\alpha}(\PP^N|H,d)$ is toroidal this locus is open and dense. By definition $\Dcal_{1,\alpha,k}(\PP^N|H,d)$ is contained in the  complement of this locus and can be expressed as a union of strata of positive codimension. Since $\Dcal_{1,\alpha,k}(\PP^N|H,d)$ has codimension $1$, it must therefore be equal to a union of logarithmic divisors.\end{proof}

In \S \ref{subsection indexing strata} we discussed a procedure for enumerating the logarithmic strata of $\VZ_{1,\alpha}(\PP^N|H,d)$, using the cones of the tropicalization $\Sigma_{1,\alpha}(\PP^N|H,d)$. Every divisorial logarithmic stratum is obtained as follows:
\begin{enumerate}
\item choose a combinatorial type $\Delta$ of a tropical map;
\item subdivide the corresponding tropical moduli space $\sigma$ to align the type;
\item choose a ray in this subdivision.
\end{enumerate}
This process contains some redundancies: upon choosing a ray in the tropical moduli space, some of the edge lengths or vertex positions may get set to $0$. This induces a generisation of the initial combinatorial type $\Delta$, given by contracting the corresponding edges of the dual graph and moving the corresponding vertices from $\RR_{>0}$ to $0$. When we speak of the combinatorial type of a stratum in $\VZ_{1,\alpha}(\PP^N|H,d)$, we will always mean this generisation. This is independent of the choice of $\Delta$, and in fact we can and will always choose $\Delta$ to coincide with the generisation.

@@

The map $\Dcal\to\Ecal$ is an isomorphism away from the exceptional centres, and the latter is contained in the locus where the core is contracted. Given an element of $\Dcal$ we can split it along the nodes $q_1,\ldots,q_r$. It is then clear that $\sqC_1$ is aligned. We claim that $\sqC$ satisfies the factorisation property if and only if $\sqC_1$ does. This immediately implies the lemma.

@@

We summarise that the contribution of each term $\lambda_\Dcal [\Dcal]$ is given by integrating over the appropriate fibre product and then multiplying the result by $\lambda_\Dcal \cdot \deg\rho = \prod_{i=1}^r m_i$.

@@

For each vertex of the target there must always be at least one vertex in the preimage which lies on the circle. This is implied by the fact that the cone $\tau$ corresponding to $\Dcal$ is a ray. If the vertex on the circle is strictly semistable -- so that in particular the collapsed map $f_B$ is constant along the corresponding component -- it can be seen that the factorisation condition is unaffected by $\Gm$-scaling the coordinate corresponding to that vertex in the moduli space of attaching data, so the map $\Dcal\to\Ecal$ has generic fibre of positive dimension and the contribution vanishes.

@@

Finally, in case the maximal contact order is $m=1$, the image tangent vectors must be linearly dependent, in which case we can deduce as above that factorisation will be satisfied independently of the choice of alignment.
\end{proof}
For the remainder of this section we use the following notation for the combinatorial type $\Delta$; we let $\sqC_0$ denotes the vertex of dual graph corresponding to the core, $\sqC_1,\ldots,\sqC_r$ the other stable vertices (genus zero and lying on the circle, by the previous lemma) and $q_1,\ldots,q_r$ the corresponding splitting nodes.

@@

The combinatorics of the flattening can be described explicitly. Consider the map on tropicalizations
\[
\widetilde{\Dcal}^{\operatorname{trop}} \to \Ecal^{\operatorname{trop}}.
\]
The tropical curves on the left inherit a central alignment from the moduli space $\widetilde{\VZ}_{1,\alpha}(\PP^N|H,d)$, which orders a certain subset of the edge lengths. The polyhedral criterion for equidimensionality is that every cone of the source maps surjectively onto a cone of the target. There is a universal logarithmic modification $\widetilde\Ecal \to \Ecal$ such that $\widetilde\Dcal \to \widetilde\Ecal$ is flat, constructed by taking the minimal subdivision of $\Ecal^{\operatorname{trop}}$ such that cones map surjectively to cones.

@@

\begin{lemma}
 For every $\Lambda_B\subseteq \{\sqC_1,\ldots,\sqC_r\}$ and $\Lambda_F=\{\sqC_1,\ldots,\sqC_r\}\setminus \Lambda_B$, $s$ vanishes along the loci where the factors corresponding to $\Lambda_B$ in $\widetilde \Ecal$ degenerate so that $f_i\colon C_i\to\PP^N$ is constant in a neighbourhood of $q_i$, and the fibre coordinates $(\lambda_j)_{j\in\Lambda_F}$ along $p\colon\mathcal P\to\widetilde \Ecal$ are set to be $0$.
\end{lemma}
Call these loci $\Xi_\Lambda$ for $\Lambda=(\Lambda_B,\Lambda_F)$ (compare with \cite[\S 3.2]{VZ} for an analogous phenomenon in the non-relative case); clearly, they may be in excess with respect to the expected codimension of $V(s)$, that is $N-1$. These loci are logarithmic substrata; therefore, there exists a logarithmic blow-up that principalises them. We claim which $\widetilde\Dcal$ is (possibly a refinement of) such a blow-up. 
\begin{lemma}
 The pullback of $\Xi_\Lambda$ is principal in $\widetilde\Dcal$.
\end{lemma}
\begin{proof}We analyse the ideal of $\Xi_\Lambda$: it is generated by
 \begin{enumerate}
  \item equations on $\widetilde\Ecal$ cutting the locus where $C_i$ degenerates so that $f_i$ is constant on a neighbourhood of $q_i$ for $i\in \Lambda_B$: these correspond to the tropical functions $\delta_i-\lambda(q_i)$ where $\lambda$ is the distance function to the core;
  \item equations on the fibre of $p$ for the linear subspace corresponding to $i\in \Lambda_F$: tropically, these correspond to the $\ell_i$ discussed in Lemma \ref{lem:generic_proj_bundle}.
 \end{enumerate}
By construction, on $\widetilde \Dcal$ it is always possible to tell which of these tropical lengths is the shortest (the circle we have drawn on $\Xi_\Lambda$ is degenerate -- it passes through no vertex of positive horizontal degree -- therefore we need to decide for a strictly larger circle in order to lift to $\widetilde D$), which corresponds to principalising the ideal (see Figure \ref{fig:principalisation}).
\end{proof}

@@

%Draw ldots
\draw [color=blue] (4,0.5) node[right]{$\ldots$};
\end{tikzpicture}
\end{center}
\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
with all other marked points contained on $\sqC_1$. This is isomorphic to $\VZ_{1,\alpha}(X|Y,\beta)$ and when we restrict $\tilde\gamma$ to this locus we obtain precisely the class $\gamma$ which we started with. Thus the left-hand side of \eqref{step 1 recursion} may be written as
\begin{equation*} (\alpha_1-1) I + \fgt_y^\st \left( (\alpha_1-1)\psi_1 + \ev_1^\st Y\right)  \tilde\gamma \cap [\VZ_{1,(\alpha-e_1)\cup(1)}(X|Y,\beta)]\end{equation*}
where $I$ is the invariant we are trying to compute. The second term is a forgetful relative invariant with the same degree and number of true markings (since $y$ is fictitious), and strictly smaller true tangency; hence it is recursively known. We now examine the right-hand side of \eqref{step 1 recursion}. Recall that all of the genus zero data has already been computed, so we only need to focus on the genus one pieces. First consider the type I loci. The genus one piece has strictly smaller degree (and hence is recursively known) except in the following situation (with some stable distribution of the remaining markings):
\begin{center}
\begin{tikzpicture}[scale=1]
\draw (0,0) to (2,0);
\draw [fill=white] (0,0) circle[radius=3pt];
\draw (0,-0.1) node[below]{\small$d$};
\draw (0,0) node[above]{\small$\sqC_1$};
\draw [fill=black] (2,0) circle[radius=3pt];
\draw (2,-0.1) node[below]{\small$0$};
\draw (2,0) node[above]{\small$\sqC_0$};
\draw [->] (2,0) -- (3,0);
\draw (2.9,0) node[right]{$x_1$};

@@

We will make use of a recursion formula for rubber spaces analogous to the recursion formula for relative spaces used in Steps 1--3. Following \cite{EKatz}, there is a line bundle $L_y^{\not\in \operatorname{bot}}$ on $\VZ^{\leftrightarrow}_{1,\alpha\cup (0)}(\mathbb P|Y_0+Y_\infty,\beta)$, together with a section $s_y^{\not\in\operatorname{bot}}$ whose vanishing locus consists of the locus $\Dcal(y)$ of rubber maps where $y$ is not mapped into the bottom level of the expanded target. As in \S \ref{Section Gathmann line bundle}, we can give a logarithmic interpretation of this: it corresponds to the piecewise-linear function on the tropical moduli space which associates, to every rubber tropical map, the distance between $\varphi(\sqC_y)$ and the leftmost vertex of the tropical target, where $\sqC_y$ is the vertex of the source curve containing the flag corresponding to $y$. Using this tropical description, we may easily calculate the vanishing orders of $s_y^{\not\in\operatorname{bot}}$ along the various components of $\Dcal(y)$, and show that $\cchern_1(L_y^{\not\in\operatorname{bot}}) = \Psi_0 - \ev_y^\st Y$ (see also \cite{EKatzLB}, where similar results in the non-reduced setting are obtained, using different methods). From this, we obtain a rubber recursion formula
